Red Wine Explorations By Xin Ding
========================================================
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(scales)
library(GGally)
library(reshape2)
library(memisc)
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
redwines<- read.csv("wineQualityReds.csv")
redwines<- subset(redwines, select = -c(X))
```

This report explores a dataset containing qualities and attributes for 
1599 red wines.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(redwines)
str(redwines)
summary(redwines)
```

Our dataset consists of 12 variables, with almost 1599 observations.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = quality)) +
  geom_histogram(stat = "count")
```

Wine qulities is skewed to the right with most qualities condense on level 5 and 
level 6.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = fixed.acidity)) +
  geom_histogram(binwidth = 0.1)+
  labs(x ="fixed acidity (tartaric acid - g / dm^3)")
```

The level of non volatile acid in red wine is between 4.6 to 15.9g per liter 
wine. Most wines have it between 7 g/L to 9g/L.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = volatile.acidity)) +
  geom_histogram(binwidth = 0.01)+
  labs(x ="volatile acidity (acetic acid - g / dm^3)")
summary(redwines$volatile.acidity)
```

Volatile acidity represents the amount of acetic acid in wine which affects the 
sourness in wine taste. This varible seems to have a close relationship with 
wine quality.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = citric.acid)) +
  geom_histogram(binwidth = 0.01)+
  labs(x ="citric acid (g / dm^3)")
```
```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = citric.acid)) +
  geom_histogram(binwidth = 0.01)+
  labs(x ="citric acid (g / dm^3)")+
  xlim(0, quantile(redwines$citric.acid, .95))
```

The amount of citric acid brings the freshness and taste in red wines. A number 
130 wines contain none citric acid is questionable, I adjusted the 
confidence level of the data to clean out some outliers.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = pH)) +
  geom_histogram()+
  labs(x ="volatile acidity (acetic acid - g / dm^3)")
```

pH value describe the acidity in red wine from 0 (acidic) to 14 basic, most 
wines are between 3 and 4 on the scale.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = residual.sugar)) +
  geom_histogram(binwidth = 0.1)+
  labs(x ="residual sugar (g / dm^3)")
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.1)+
  labs(x ="alcohol (% by volume)")
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = density)) +
  geom_histogram()+
  labs(x ="density (g / cm^3)")
```

Wine body so called wine density is depending on the on the percent alcohol and
sugar content.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = chlorides)) +
  geom_histogram(binwidth = 0.001)+
  labs(x ="chlorides (sodium chloride - g / dm^3)")
```



```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = free.sulfur.dioxide)) +
  geom_histogram(binwidth = 1)+
  labs(x ="free sulfur dioxide (mg / dm^3)")
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = total.sulfur.dioxide)) +
  geom_histogram(binwidth = 1)+
  labs(x ="total sulfur dioxide (mg / dm^3)")
```

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = redwines, aes(x = sulphates)) +
  geom_histogram(binwidth = 0.01)+
  labs(x ="sulphates (potassium sulphate - g / dm3)")
```

The antioxident composition inside wines, existing as in forms of sulfur dioxide
if free concentration more than 50 ppm, human nose would sense it, as well as 
being tasted in the wine.

# Univariate Analysis
<!--
> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!
-->

### What is the structure of your dataset?
There are 1599 red wines in the dataset with 12 features (fixed acidity, 
volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, 
total sulfur dioxide, density, pH, sulphates, alcohol and quality). The variable 
quality is output variable with scale from 1 to 10, which larger in level 
number, better in wine quality.

### What is/are the main feature(s) of interest in your dataset?
According to descirption, I can allege that volatile acidity plays a significant
role in the quality. Meanwhile, density and sulfur elements are expected to 
influence the taste at a certain degree.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
Alcohol and pH values

### Did you create any new variables from existing variables in the dataset?
Yes, I created conditional means for volatile acidity and alcohol.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Nearly 130 wines contain none citric acid is questionable, I adjusted the 
confidence level of the data to clean out some outliers.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}
correlation <-cor(redwines)
round(correlation, 2)
```

The quality of red wine tends to correlate with volatile acidity and alcohol 
content percentage. Respectively, volatile acidity and alcohol percentage
correlates with citric acid and density. However, there is not a single variable
strongly correlated to each other. One plausible reason is too few observations.

```{r echo=FALSE}
sub_redwine <- subset(redwines, select = c(volatile.acidity, citric.acid, density, alcohol, quality))
ggpairs(sub_redwine,lower = list(
          continuous = "smooth",
          combo = "facetdensity",
          mapping = aes(color = quality)),
        upper = list(continuous = wrap("cor", size = 3, hjust=0.8)))
```

After visualizing all the relatively correlated variables, due to low 
correlation between the variables, wine quality is slightly positive 
proportional to alcohol, and decreases as volatile acid increases. 

To better visualize the relationship between quality and other variables, I 
decide to group the dataset by quality, and take conditional mean of the varible
for the plot.

```{r echo=FALSE,message=FALSE, warning=FALSE}
modified_redwines <- redwines
modified_redwines$quality <- ordered(modified_redwines$quality, levels = c(1,2,3,4,5,6,7,8))
ph_group <- summarise(group_by(modified_redwines, quality),
                              fixed_acid.mean = mean(fixed.acidity),
                              ph = mean(pH),
                              n = n())
ggplot(data = modified_redwines, aes(x = quality , y =alcohol)) +
  geom_boxplot()+
  labs(y ="alcohol (% by volume)")+
  ggtitle("Alcohol VS Quality")
```

Once alcohol perentatge pasts approximately 10.5 pecent, the quality of wine 
tends to escalate linearly.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = modified_redwines, aes(x = quality , y =volatile.acidity)) +
  geom_boxplot()+
  labs(y ="volatile acidity (acetic acid - g / dm^3)")+
  ggtitle("Volatile acidity VS Quality")
```

Sour tasting wines receive low grades.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
Quality is almost proportional to alcohol percent, but reciprocal to the amount 
of volatile acid.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
Citric acid correlated negetively with volatile acid indeed surprised me. I was
expecting that citric acid helps to keep evaporative acid in wine.

### What was the strongest relationship you found?
From the cor tables, it said the strongest relationship was -.69 between pH and fixed 
acidity. 

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}

ggplot(data = modified_redwines, aes(x= pH, y= fixed.acidity))+
  geom_smooth(aes(color=quality)) +
  scale_colour_brewer(palette = "Blues")+
  labs(y ="fixed acidity (g / dm^3)")+
  ggtitle("Fixed acidity VS pH by Quality")
```

For better visulazing the relationship between fixed.acidity and pH respected to 
different qualities, I transformed the quality into ordered factor. As you can
see, a deceasing in fixed acidity in better quality wine pH drop faster for
pH greater than 3.5.

```{r echo=FALSE,message=FALSE, warning=FALSE}
ggplot(data = modified_redwines, aes(x= pH, y= volatile.acidity))+
  geom_smooth(aes(color=quality))+
  scale_colour_brewer(palette = "Blues")+
  labs(y ="volatile acidity (acetic acid - g / dm^3)")+
  ggtitle("Volatile acidity VS pH by Quality")
```

```{r echo=FALSE,message=FALSE, warning=FALSE }
ggplot(data = modified_redwines, aes(x= density, y= alcohol))+
  geom_point()+
  stat_smooth(method = "lm", col = "red")+
  facet_wrap(~quality) +
  ggtitle("facet wrap by quality")+
  labs(y ="alcohol (% by volume)")
```

Alcohol reduces the density, on the contrary that residual sugar from 
fermentation make a heavy wine body. The balance between the sugar and the 
alcohol depends how wine makers control the process of fermentation and
quality of the grapes.

```{r echo=FALSE,message=FALSE, warning=FALSE}
m1 <- lm(I(quality) ~ I(volatile.acidity), data = redwines)
m2 <- update(m1, ~ . + alcohol)
m3 <- update(m1, ~ . + total.sulfur.dioxide)
m4 <- update(m1, ~ . + sulphates)
m5 <- update(m1, ~ . + citric.acid)
mtable(m1,m2,m3,m4,m5)
```

The variables in this linear model can account for 20% or 30% at most of the 
variance in the quality of wines. Not a great model.
# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
A deceasing in fixed acidity in better quality wine pH drop faster for
pH greater than 3.5. 

### Were there any interesting or surprising interactions between features?
Wine body is one of the major criteria of wine assessment, I discovered that
alcohol reduces the density, on the contrary that residual sugar from 
fermentation makes wine heavy.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.
Yes, I have to admit this linear model should not be used as a prediction model 
for wine quality.The variables in this linear model can account for 20% or 30% at 
most of the variance in the quality of wines. Due to the low correlation between
variables and output variable, we will not be able to build a strong linear 
regression model base on the data given.
------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
ggplot(data = modified_redwines, aes(x= density, y= alcohol))+
  geom_point()+
  stat_smooth(method = "lm", col = "red")+
  facet_wrap(~quality) +
  ggtitle("facet wrap by quality")+
  labs(y ="alcohol (% by volume)")
```

### Description One
One concept I need to introduce before we go further is wine body. It is an
anlysis a way when the wine connoisseur decribes how wine feels in mouth after 
the taste. 

Three categories: light(<12.5% alcohol), medium(12.5% to 13.5% alcohol) and 
full(>13.5% alcohol). The fact is that, nowadays, wine tasters prefer full body 
wines to light or medium wines. Presumably, depending on people's taste, a 
better quality wine is heavier in body.

However, we have been told that body has been defined as the more alcohol wine
contains, it becomes more viscous so that it becomes heavier. 

This concept is not well supported by our analysis. We can see that the density 
increases when there is less alcohol. So, can people really tell how good a wine 
by their tongues?

### Plot Two
```{r echo=FALSE,message=FALSE, warning=FALSE, Plot_Two}
ggplot(data = modified_redwines, aes(x = quality , y =alcohol)) +
  geom_boxplot()+
  labs(y ="alcohol (% by volume)")+
  ggtitle("Alcohol VS Quality")
```

### Description Two
The concept has been proved wrong in the previous description. But the fact that
people prefer full bodied wine seems to appear to be true in the graph above.
We can deduce that somehow people does taste more flavor in wines with more
alochol in it. Another theory saying that more alcohol would extract other 
elements and molicules in grapes might explain a little, or they just get pleased
with more alcohol.

### Plot Three
```{r echo=FALSE,message=FALSE, warning=FALSE, Plot_Three}
ggplot(data = modified_redwines, aes(x = quality , y =volatile.acidity)) +
  geom_boxplot()+
  labs(y ="volatile acidity (acetic acid - g / dm^3)")+
  ggtitle("Volatile acidity VS Quality")
```

### Description Three
Another interesting discover here is that people does not like volatile acid. It 
stimilates our palate with sourness. You don't want a vinegar taste wine.


# Reflection
The red wines data set contains information on 1599 wines across twelve 
variables from around 2009. I started by understanding the individual variables 
and exploring interesting questions by making observations on plots. I did make
a linear model to predict the quality of wine, but it was not reliable.

Nothing appeared to be clear just by ploting them on the graph at first. 
After using the correlation table, I found that the variables were not strongly 
related to each other, or in other words, data set we had could not lead to an
accurate analysis. By picking the most correlated varibles, I plotted several 
graphs to tell a story. According to cor table, the highest correlations were 
between quality and alcohol, between quality and volatile acid are 0.48 and 
-0.39 respectively.They were still fairly weak, but we knew where to start now. 
I drew boxplots using quality versus alcohol and volatile acid. The conclusion 
was that wine quality increases when there is more alcohol and less aetic acid 
in it. Finally, I made a linear model based on all chemicals included 
in the wine even though I knew at the beginning it would not be a practical 
model. It was still worth a shot.

limitations of this model include data source. Given that at least 3 wine 
experts rated the quality of each wine, providing a rating between 0 and 10 , 
results could be very bias since wine tasting is highly depending on experts'
personal judgements, not the universal standard.

For future analysis on red wines, a data set includes pubilc grading through 
blind testing is recommended. Compared to the wine experts judgement, a 
generalized data is relatively unbias. As a reference, we can figure out if 
public agrees that higher score wine graded by experts tastes better.